---
title: "Course Project 1"
author: "Hamed Yazarloo"
date: "March 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/hyazarloo/Documents/GitHub/RepData_PeerAssessment1/RepData_PeerAssessment1")
```



# Loading and preprocessing the data

### Load the data
```{r}
data<-read.csv("./data/activity.csv",colClasses = c("numeric","character","numeric"))
```
### Process/transform the data into a format suitable for the analysis
```{r}
data$date <- as.Date(data$date, "%Y-%m-%d")
```
# What is mean total number of steps taken per day?
The missing values in the dataset have been ignored.
### Calculate the total number of steps taken per day
```{r}
sums<-aggregate(steps ~ date,data, sum)
```
### Make a histogram of the total number of steps taken each day
```{r}
hist(sums$steps, col = "green",main = "Histogram of total steps per day",xlab = "steps",ylab = "Frequency")
```
### Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean<-mean(sums$steps,na.rm = TRUE)
median<-median(sums$steps,na.rm = TRUE)
```
# What is the average daily activity pattern?

### A time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
average_of_intervals<-tapply(data$steps, data$interval, mean, na.rm = TRUE)
names(average_of_intervals)<-sort(unique(data$interval))
plot(average_of_intervals~names(average_of_intervals),type = "l", col="red",xlab = "Interval",ylab = "Average", main = "Average of steps vs Intervals")
```
### The 5-minute interval, on average across all the days in the dataset, containing the maximum number of steps:
```{r}
names(average_of_intervals)[average_of_intervals==max(average_of_intervals)]
```
# Imputing missing values

### The total number of missing values in the dataset
```{r}
sum(is.na(data[, 1]))
```
NAs have been replaced by the average value for each interval.
A new dataset is created that is equal to the original dataset but with the missing data filled in.
```{r}
data2<-data
data2[, 1][is.na(data2[,1])]<-round(average_of_intervals[as.character(data2[, 3][is.na(data2[,1])])])
```
### A histogram of the total number of steps taken each day:
```{r}
sums2<-aggregate(data2[, 1], list(data2$date), sum)
names(sums2)<-c("day","total")
hist(sums2$total)
```
### The mean and median total number of steps taken per day:
```{r}
mean2<-mean(sums2$total,na.rm = TRUE)
median2<-median(sums2$total,na.rm = TRUE)
```
# Differences in activity patterns between weekdays and weekends

### A new factor variable in the dataset is created with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
data2[,2]<-as.Date(data2[,2])

library(lubridate)
library(dplyr)
data2<-mutate(data2,weekday=wday(date))
data2$weekday[data2$weekday %in% c(7,1)]<-"weekend"
data2$weekday[data2$weekday %in% c(2,3,4,5,6)]<-"weekday"
data2$weekday<-factor(data2$weekday)
```
### A panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
library(plyr)
data2$date<-NULL
weekday_factor<-ddply(data2,.(interval,weekday),summarize, mean = round(mean(steps), 2))
library(lattice)
xyplot(mean ~ interval | weekday, weekday_factor,
       type = "l", layout = c(1, 2), xlab = "Intervals", ylab = "Count of steps")
```




